---
title: "Assignment-3"
author: "Ajay shadam"
date: "2023-10-11"
output:
  html_document: default
  pdf_document: default
---

```{r}
## Load required libraries
library(tidyverse)
library(reshape)
library(reshape2)
library(caret)
library(e1071)

## Load the data csv file
Bank_data <- read.csv("C:/Users/Ajay Reddy/Downloads/UniversalBank-1.csv")

head(Bank_data)
tail(Bank_data)
```

```{r}
## specifying names for columns

colnames(Bank_data)

#changing information into components using classifications.


Bank_data$`Personal Loan` = as.factor(Bank_data$Personal.Loan)
Bank_data$Online = as.factor(Bank_data$Online)
Bank_data$CreditCard = as.factor(Bank_data$CreditCard)
```

```{r}
#partitioning the data into half, with 40% going into the set used for validation and 60% flowing into the initial training set


set.seed(456)

# 60 % training data and 40% Validation data respectively
Bank_data.train.data <- sample(row.names(Bank_data), 0.6*dim(Bank_data)[1])
Bank_data.valid.data <- setdiff(row.names(Bank_data), Bank_data.train.data)

# inserting validation and training data, respectively, through a dataframe from a dataset


Bank_data.train <- Bank_data[Bank_data.train.data, ]
Bank_data.valid <- Bank_data[Bank_data.valid.data, ]

# Duplicating the data frame UB.train
train <- Bank_data[Bank_data.train.data, ]
# Duplicating the data frame UB.valid
valid = Bank_data[Bank_data.train.data,]

print(train)
print(valid)
```

## A. Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().

```{r}
## merge data and data


melt_datafram = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")

# Data which has melted is cast


data_casted <- dcast(melt_datafram, CreditCard + Personal.Loan ~ value, fun.aggregate = length)

# Credit card and ID, and personal loan are listed in casting column number 14, in accordance.

data_casted[,c(1,2,3,14)]
```

#B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

```{r}
#The pivot table demonstrates that the value for CC is whereas the value under Loan is 89. 

Bank_data.Loan.CC1 <- 89/3000

# which is 29 %.
Bank_data.Loan.CC1
```
## C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}
 ## converting the train data set into an extended format, using "Online" as an attribute to be melted and "Personal.Loan" as an identification

melt1 = melt(train,id=c("Personal.Loan"),variable = "Online")

 ## creating an extended format out of the train data frame, using "CreditCard" as an identifier and "Online" as an attribute that needs to have been melted

melt2 = melt(train,id=c("CreditCard"),variable = "Online")
```

```{r}
# The casting process Online values and personal loans

# Understanding that Online and Personal Loans are binary (0 or 1)

cast1 =dcast(melt1,`Personal.Loan`~Online) 

# The casting process Online values and personal loans

cast2=dcast(melt2,CreditCard~Online)
```

```{r}
#illustrates the total amount of personal financing available online.

Bank_data.Loanonline=cast1[,c(1,13)]
Bank_data.LoanCC = cast2[,c(1,14)]
Bank_data.Loanonline
```
```{r}

# exhibits the quantity of credit card transactions in relation to the internet.

Bank_data.LoanCC

```
## D. Compute the following quantities [P (A | B) means “the probability of A given B”]:
#1.P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) #2.P(Online=1|Loan=1) 
#3.P (Loan = 1) (the proportion of loan acceptors)
#4.P(CC=1|Loan=0)
#5.P(Online=1|Loan=0)
#6.P(Loan=0)

```{r}
# constructing a pivot tables with personal loans illustrated by the columns 14 and 10.


table(train[,c(14,10)])
```
```{r}
# Implementing a pivot table to display online and personal loan columns 13 and 10.



table(train[,c(13,10)])
```

```{r}
 # A pivoting table for personal loans These are, from training, 2725 and 275, correspondingly.


table(train[,c(10)])
```
```{r}
## By referencing the previously mentioned p, we may get the values for CC = 1 and Loan = 1.


Bank_data.CCBank_data.Loan1 = 89/(89+200)
Bank_data.CCBank_data.Loan1
```
2. P(Online=1|Loan=1)

```{r}
##The table of pivot points above UB.ONUB.Loan1 presents us with the values we need for online = 1 and loan = 1.


Bank_data.ONBank_data.Loan1 =177/(177+112)
```


3. P(Loan = 1)

```{r}
 # Using the pivot point table above, we are able to determine that Loan = 1.


Bank_data.Loan1 =289/(289+2711)
Bank_data.Loan1
```


4. P(CC=1|Loan=0)

```{r}

#We have the ability to establish the values for CC = 1 and Loan = 0 by using the pivot table described above.


Bank_data.CCLoan.01= 794/(794+1917) 
Bank_data.CCLoan.01
```

5. A(Online=1|Loan=0)

```{r}
# From the table of pivot points above, we can determine the following values: online = 1 and loan = 0.

Bank_data.ON1.L0= 1665/(1665+1046)
Bank_data.ON1.L0
```

6. P(Loan=0)

```{r}
# By using the pivot table that was created above, we can find the values for Loan = 0.

Bank_data.Loan0= 2711/(2711+289)
Bank_data.Loan0
```
##E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1, Online = 1).

```{r}
# Given probabilities
A_CC_data <- 0.096
online_givendata_fram <- 0.833
loan.1 <- 0.0125

# Calculate Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)
#Bank_data_Naivebayes <- (A_CC_data * online_givendata_fram * loan.1) / (P_CC * P_Online)

Bank_data_Naivebayes <- (A_CC_data)*(online_givendata_fram)*(loan.1)
Bank_data_Naivebayes

```

## F. Compare this value with the one obtained from the pivot table in (b). Which is a more accurate estimate?

```{r}
# Possibility of Naive Bayes (from computation in E).

prob.naivebayes <- 0.0009996

# Possibility Pivot Table (from part B)

pivot.prob_tab <- 0.02966667

#After evaluating the odds, display a message indicating which is most likely to take place.

if (prob.naivebayes > pivot.prob_tab) {
  message("Naive Bayes Probability is more accurate: ", prob.naivebayes)
} else if (prob.naivebayes < pivot.prob_tab) {
  message("Pivot Table Probability is more accurate: ", pivot.prob_tab)
} else {
  message("Both probabilities are the same: ", prob.naivebayes)
}

```
##The table with the pivot probability (0.02966667) is thought to be greater in precision than the Naive Bayes possibility based on the comparison.



## G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}

names(Bank_data)

names(Bank_data.train)

# Enter the appropriate ones in the training columns.

Bank_data.train <- Bank_data[, c("CreditCard", "Online", "Personal Loan")]

# Swap out the spaces simply renaming the columns

colnames(Bank_data.train) <- c("CreditCard", "Online", "PersonalLoan")

# Convert "CreditCard" and "Online" become factors which have the proper magnitude.

Bank_data.train$CreditCard <- factor(Bank_data.train$CreditCard, levels = c(0, 1), labels = c("No", "Yes"))
Bank_data.train$Online <- factor(Bank_data.train$Online, levels = c(0, 1), labels = c("No", "Yes"))
```


# Print the probability
print("Probability of Loan = 1 given CC = 1 and Online = 1:")
print(prob_loan_2)

```


